---
title: "Basic Inferential Data Analysis Instructions"
author: "Michael Steiner"
date: "October 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data
####Description
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

####Format
A data frame with 60 observations on 3 variables.
[,1] len numeric Tooth length
[,2] supp factor Supplement type (VC or OJ).
[,3] dose numeric Dose in milligrams.
```{r set_known}
library(datasets)
data("ToothGrowth")

```
##Summary
```{r summary}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
```
```{r sumary_plot}
oj <- ToothGrowth[ToothGrowth$supp %in% c('OJ'),]
vc <- ToothGrowth[ToothGrowth$supp %in% c('VC'),]
plot(as.numeric(oj$dose), oj$len, col = 1, pch = 1, xlab = 'Dose (mg/day)', ylab = 'Length', main = '')
points(as.numeric(vc$dose), vc$len, col = 2, pch = 2)
legend('bottomright', c("OJ", "VC") ,bty = "n", pch = c(1,2), col = c(col = 1, col = 2))
```

Preliminary analysis shows a general trend of longer teeth with increasing dosage of both Orange Juice and Vitamin C. 

##Confidence intervals
```{r test}
t.test(len ~ supp, data = ToothGrowth)

```
The p-value is larger than the 5% significance level so the null hypthosis cannot be rejected.
```{r test2}
data("ToothGrowth")
t.test(ToothGrowth$len, ToothGrowth$dose)
```
When dosage is accounted for, the p-value becomes approximatly 0 and is smaller than the significance level. 

##Conclusions
The t-test results can be used to conclude that there is not a difference in tooth length based on the dosage of either VC or OJ but that the actual suppliment does not have a significant effect on tooth growth. 
